A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-28 latex/paper.tex
Modified the engine sub-section

\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
engine. It provides in-memory computing, data locality, and lazy
evaluation. Spark's main data structure, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
elements. 
Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
performance features. 
Spark's core is written in Scala, but it also has an R and a Python API
called PySpark. We used PySpark, given the popularity of Python in
neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler, a
process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches
tasks to them. The Spark standalone scheduler uses a LIFO
(Last-In-First-Out) task scheduling policy \TG{we should describe it
briefly and refer to the paper}. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated  process
on the cluster, and cluster mode, where the driver runs in a worker. We
used client mode as PySpark is not available in cluster mode. We used
Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. \TG{A few more
introductory sentences should be added, following Spark's model.}
\TG{Explain workers}
Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler.
Dask Distributed also has a LIFO policy. In theory, it should speed-up the
pipeline execution by finishing a branch of task computations before
starting new ones \TG{This is meant to reduce the memory footprint of the
application}. \TG{Here we should add a few comparison sentences between
Spark's and Dask's schedulers.} Dask has five main APIs and data structures: Bag,
Array, Futures, Delayed and DataFrame. We used them all
except DataFrame as it is not adapted to the type of data we
used. All APIs provide in-memory computing, data locality, and lazy
evaluation, however, none of them has fault tolerance \TG{Check fault
tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a
programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}.
An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for
the processing of large arrays. It provides a distributed clone of the
popular
NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
tasks that do not fit in the Array, DataFrame or Bag APIs. 
Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
Delayed as they support arbitrary tasks, but they operate in real-time rather
than lazily.

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8
 threads. We used Dask's default configuration that uses all the available
 memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
 configured Spark to use 1 executor per worker and 25~GB of memory per
 executor, to leave 5~GB for overhead. \TG{Check spill to disk}

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Experiment 1: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
the y-axis and the number of workers on the x-axis; the color represents the
different frameworks. Overall, there is no substantial makespan difference between
the frameworks. Also, the increase in performance is not proportional to the number
of workers; performance even decreases when we reach 8 workers.

@@

% blocks
\subsection{Experiment 1: Number of blocks}
In Figure~\ref{fig:inc_block}, Spark is not compared for 30 blocks. This is because
Spark has a 2GB limitation in the task size it can compute.

@@

%% Iterations
\subsection{Experiment 1: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
number of iterations. Spark and the Dask Bag API are equivalent in terms of
performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
it has a similar makespan to Spark and Dask Bag when the number of iterations
increases. Overall, Dask Futures seems to be faster but the difference is not
substantial.

@@

%% Sleep time
\subsection{Experiment 1: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
different sleep delays. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.

>>>>

\newcommand{\TG}[1]{\color{cyan}From Tristan: #1 \color{black}}
\newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's other popular API, is
an overlay of RDDs with similar performance features. Spark's core is written in
Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
the popularity of Python in neuroscience, although the serialization cost of Python
objects to Scala might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches jobs to them. A job is divided
into stages to be processed onto the workers. Each stage's operation is represented
as a high-level task in the computation graph. The Spark standalone scheduler uses a
FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
increasing the page hit rate. As our applications only have one job, the LIFO
(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
phrasing} \TG{refer to the paper (\MD{Not sure which one})}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as PySpark is not available in cluster mode. We used Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
in the scientific Python ecosystem. Dask has five main APIs and data structures:
Array, Bag, DataFrame, Delayed, and Futures. We used them all except DataFrame as it
is not adapted to the type of data we used. Like Spark, all Dask APIs provide
in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
lineage. Dask's core is purely in Python and bypass most of the
\href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
We used Dask due to the common use of the scientific Python ecosystem in
neuroscience. Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. In Dask, operations generate multiple
tiny tasks in the computation graph allowing an easier representation of complex
algorithms. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler, although \href{https://github.com/dask/dask-yarn}{Dask-Yarn}
and \href{https://github.com/mrocklin/dask-mesos}{Dask-Mesos} is also available. In
the Dask Distributed scheduler, a process called \textit{dask-scheduler}
administrates the resources provided by \textit{workers} in the cluster. The
scheduler receives jobs from clients and assigns tasks to available workers. Like
Spark scheduler, Dask Distributed also use a LIFO policy, it is meant to reduce the
memory footprint of the application by finishing the computation of a branch in the
Dask graph before starting new ones. We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a programming
abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
parallel composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8
 threads. We used Dask's default configuration that uses all the available
 memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
 configured Spark to use 1 executor per worker and 25~GB of memory per
 executor, to leave 5~GB for off-heap. \TG{Check spill to disk}

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
the y-axis and the number of workers on the x-axis; the color represents the
different frameworks. Overall, there is no substantial makespan difference between
the frameworks. Also, the increase in performance is not proportional to the number
of workers; performance even decreases when we reach 8 workers.

@@

% blocks
\subsection{Incrementation: Number of blocks}
In Figure~\ref{fig:inc_block}, Spark is not compared for 30 blocks. This is because
Spark has a 2GB limitation in the task size it can compute.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
number of iterations. Spark and the Dask Bag API are equivalent in terms of
performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
it has a similar makespan to Spark and Dask Bag when the number of iterations
increases. Overall, Dask Futures seems to be faster but the difference is not
substantial.

@@

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
different sleep delays. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.
